#Tools for visualising amplicon data
ampvis is an R-package to visualise amplicon data. more more more

##Installing ampvis2
First, install [R](https://mirrors.dotsrc.org/cran/) (and for Windows also [RTools](https://mirrors.dotsrc.org/cran/bin/windows/Rtools/)) and [RStudio](https://www.rstudio.com/products/rstudio/download/#download). Then open RStudio and install the following required packages from the console. It can be a good idea to run RStudio as administrator, so the packages are available to all users on the computer. To run all commands at once copy them into a new R script and hit CTRL/CMD + SHIFT + ENTER. If you get asked to update old packages, just hit a.
```{r, message = FALSE, warning = FALSE, eval=FALSE}
install.packages("devtools")
source("http://bioconductor.org/biocLite.R")
biocLite("Biostrings")
biocLite("DESeq2")
biocLite("munsell")
biocLite("stringi")
biocLite("chron")
biocLite("permute")
biocLite("assertthat")
biocLite("rmarkdown")
biocLite("reshape")
```

Now ampvis2 can be installed from github. If you get an error about a missing package, just use `install.packages("packagename")` to install it. If that doesn't work, try with `biocLite("packagename")` instead.
```{r, message = FALSE, warning = FALSE, eval = FALSE}
devtools::install_github("MadsAlbertsen/ampvis2")
library(ampvis2)
```